---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Amanullah"
date: "28 September, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r}
#loading packages
library(dplyr)
library(readr)
library(magrittr)
library(tidyverse)
library(tidyr)
```

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

## Exercises

### Exercise 1

The `the_tate` dataset has 3336 unique artists who worked from 1545 to 2012. The works were acquired between the years 1823 and 2013.

```{r}
the_tate |> dplyr::summarize(
  artist=length(unique(artist)),
   minYear = min(year, na.rm = TRUE),  
    maxYear = max(year, na.rm = TRUE), 
    minacquisitionYear = min(acquisitionYear, na.rm = TRUE), 
    maxacquisitionYear = max(acquisitionYear, na.rm = TRUE),
)
```

```{r}
DataExplorer::introduce(the_tate)
```

### Exercise 2

How number of works with missing dates is 5397.

The number of artists whose works have missing dates is 461.

It would require resolving missing year data for only 11 artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as MAR.

```{r}
DataExplorer::introduce(the_tate)
```

```{r}
the_tate |> DataExplorer::plot_missing()
```

```{r}
library(tibble)
library(dplyr)

DataExplorer::introduce(the_tate)
DataExplorer::introduce(the_tate_artists)
DataExplorer::plot_missing(the_tate)
DataExplorer::plot_missing(the_tate_artists)

works_with_missing_year <- sum(is.na(the_tate$year))
print(works_with_missing_year)

# Filtering the 'the_tate' dataset to get rows with missing 'year'
the_tate|>dplyr::filter(is.na(year))|>dplyr::distinct(artist)
artists_with_missing_date <- the_tate |>
  filter(is.na(year)) |>
  distinct(artist) |>
  nrow()
print(artists_with_missing_date)

#Filtering the tate_dataset to get rows with missing year values
artist_missing_year_count <- the_tate|>dplyr::filter(is.na(year)) |> 
  group_by(artist) |>
  summarise(missing_year = sum(is.na(year))) |> 
  arrange(desc(missing_year)) |> 
  
  as_tibble()

print(artist_missing_year_count)

artists_with_missing_year <- nrow(artist_missing_year_count)

print(artists_with_missing_year)

artist_missing_year_count <- artist_missing_year_count |> 

# Calculating the percentage of missing 'year' values for each artist
  mutate(percentage = (missing_year / works_with_missing_year) * 100)

artist_missing_year_count <- artist_missing_year_count |>
  mutate(cumulative_percentage = cumsum(percentage)) 

artists_to_resolve_50_percent <- min(which(artist_missing_year_count$cumulative_percentage >= 50))

print(artists_to_resolve_50_percent)
```

### Exercise 3

The artist with the most works in the Tate collection is Turner, Joseph Mallord William.

The artist with the tenth-most works in the Tate collection is Warhol, Andy

```{r}
the_tate |> dplyr::group_by(artist)|>
  dplyr::mutate(n=n())|>
  dplyr::select(artist,n)|>
  dplyr::arrange(desc(n))|>
  dplyr::distinct()|>
  dplyr::ungroup()|>
  dplyr::slice(c(1,10))
```

### Exercise 4

The artist with the greatest number of works in the Tate collection represent 56.92% of the total number of works

```{r}
the_tate |> dplyr::group_by(artist)|>
  dplyr::mutate(n=n())|>
  dplyr::select(artist,n)|>
  dplyr::arrange(desc(n))|>
  dplyr::distinct()|>
  dplyr::ungroup()|>
  dplyr::mutate(
    total=sum(n),
    pct=n/total*100
  )
```

### Exercise 5

There are 23705 duplicate artist-title pairs

```{r}
total_works <- nrow(the_tate)
total_rows <- total_works


distinct_artist_title_pair <- the_tate |> select(artist,title) |> distinct()

distinct_count <- nrow(distinct_artist_title_pair)

print(total_rows) 

print(distinct_count)  

duplicated_count <- total_rows - distinct_count
print(duplicated_count)
```

### Exercise 6

The artist with the largest work in the tate collection is Therrien, Robert

The artist with the smallest work in the collection is Mesens, E.L.T. The smallest work has area 237 \$\\tex

```{r}
the_tate <- the_tate |>
  mutate(area_cmsquare= width * height)

selected_artworks <- the_tate |>
  select(artist, title, area_cmsquare) |>
  drop_na()

ordered_artworks <- selected_artworks |>
  arrange(area_cmsquare)


largest_artwork <- ordered_artworks |>
  slice_tail(n = 1)


smallest_artwork <- ordered_artworks |>
  slice_head(n = 1)

print(largest_artwork)
print(smallest_artwork)
```

### Exercise 7

...

```{r}
library(dplyr)

gender_grouped<- the_tate |>
  dplyr::left_join(the_tate_artists, by = c("artist" = "name")) |>   tidyr::drop_na(gender) |> group_by(gender) 


print(gender_grouped) 
```

### Exercise 8

The annual return in the SPX price in 2020 was 13.98%.

The corresponding price volatility was 34.70%.

```{r}
data <- read_csv("data/SPX_HistoricalData_1692322132002.csv") 
print(data)

data <- data |>
  mutate(Year = lubridate::year(as.Date(Date, format = "%m/%d/%Y")))

data<- data |>
rename("close"=`Close/Last`)

data <- data |>
  mutate(rd = log(lead(close) / close))

data<- data |>
  mutate(vard = rd^2)


summary_data <- data |> 
  group_by(Year) |>
  summarize(
    Annual_Return = (exp(sum(rd, na.rm = TRUE)) - 1)*100, 
    Annual_StdDev = sqrt(sum(vard, na.rm = TRUE))*100,
    .groups = "drop"  
  )


print(summary_data)
```

### Exercise 9

The period volatility was 48.78%

```{r}
period_return <- prod(1 + summary_data$Annual_Return) - 1
period_volatility <- sqrt(sum(summary_data$Annual_StdDev^2))


summary_rows <- tibble::tibble(
  Year = as.character("Period"),  
  Annual_Return = period_return,
  Annual_StdDev = period_volatility
)


summary_data <- summary_data |>
  mutate(Year = as.character(Year))


summary_data <- bind_rows(summary_data, summary_rows)


print(summary_data)

```

### 
