---
title: "BSMM-lab-2"
subtitle: "BSMM 8740 Fall 2023"
author: "Amanullah"
date: "28 September, 2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r}
#loading packages
library(dplyr)
library(readr)
library(magrittr)
library(tidyverse)
```

```{r load-pkg-data}
#| message: false
the_tate <- readr::read_delim("data/the-tate-collection.csv", ";", escape_double = FALSE, trim_ws = TRUE)
the_tate_artists <- readr::read_csv("data/the-tate-artists.csv")
```

## Exercises

### Exercise 1

The `the_tate` dataset has 3336 unique artists who worked from 1545 to 2012. The works were acquired between the years 1823 and 2013.

```{r}
the_tate |> dplyr::summarize(
  artist=length(unique(artist)),
   minYear = min(year, na.rm = TRUE),  
    maxYear = max(year, na.rm = TRUE), 
    minacquisitionYear = min(acquisitionYear, na.rm = TRUE), 
    maxacquisitionYear = max(acquisitionYear, na.rm = TRUE),
)
```

```{r}
DataExplorer::introduce(the_tate)
```

### Exercise 2

How number of works with missing dates is \_\_.

The number of artists whose works have missing dates is \_\_.

It would require resolving missing year data for only \_\_ artists resolve resolve at least 50% of the missing data.

The missing year data likely to be classified as \_\_\_\_.

### Exercise 3

The artist with the most works in the Tate collection is \_\_\_.

The artist with the tenth-most works in the Tate collection is \_\_\_.

### Exercise 4

The artist with the greatest number of works in the Tate collection represent \_\_\_% of the total number of works

### Exercise 5

There are \_\_ duplicate artist-title pairs

### Exercise 6

The artist with the largest work in the tate collection is \_\_\_

The artist with the smallest work in the collection is \_\_\_. The smallest work has area \_\_\_ $\text{cm}^2$

### Exercise 7

...

### Exercise 8

The annual return in the SPX price in 2020 was \_\_\_%.

The corresponding price volatility was \_\_\_%.

### Exercise 9

The period volatility was \_\_\_.\_%

### 
